Symbolic Weighted Language Models and\\ Quantitative Parsing over Infinite Alphabets

### 2021-07-15 introduction.tex
2 last examples (VPA and parsing)

\begin{example}%[Running example]
\label{ex:running}
We illustrate our framework with a very simplified %toy
running example of \emph{music transcription}: 
a given \emph{timeline} of timestamped musical events 
from an infinite alphabet $\Sigma$ as input, 
is parsed into a structured music score. 
Input events of $\Sigma$
are pairs $\mu \ts{\tau}$, where $\mu$ is a 
MIDI pitch~\cite{Selfridge-Field97beyondMIDI}, %\cite{MIDIfile}
and $\tau \in \mathbb{Q}$ is a timestamp in seconds. 
Such inputs typically correspond
to the recording of a live performance, \eg
$I = 69\ts{0.07}, 
     71\ts{0.72}, 
     72\ts{0.91}, 
     74\ts{1.05}, 
     76\ts{1.36}, 
     77\ts{1.71}$. % over interval $[0,2[$,

>>>>

\begin{example}%[Running example]
\label{ex:running}
We illustrate our framework with a very simplified %toy
running example of \emph{music transcription}: 
a given \emph{timeline} of timestamped musical events 
from an infinite alphabet $\Sigma$ as input, 
is parsed into a structured music score. 
Input events of $\Sigma$
are pairs $\mu \ts{\tau}$, where $\mu$ is a 
MIDI pitch~\cite{Selfridge-Field97beyondMIDI}, %\cite{MIDIfile}
and $\tau \in \mathbb{Q}$ is a timestamp in seconds. 
Such inputs typically correspond
to the recording of a live performance, \eg
$I = 69\ts{0.07}, 
     71\ts{0.72}, 
     73\ts{0.91}, 
     74\ts{1.05}, 
     76\ts{1.36}, 
     77\ts{1.71}$. % over interval $[0,2[$,

### 2021-07-15 main.tex
2 last examples (VPA and parsing)

\begin{example}
... \SWVPA over the alphabet of Example~\ref{ex:nested-word}
that expresses a weight related to the music notation,
more precisely to its structural complexity: 
given a set of equivalent representations, we aim at choosing the simpler one. 
...

*** SUPPR ***
For instance
\includegraphics[scale=0.20]{pictures/score5.png} will be favored
on \includegraphics[scale=0.20]{pictures/score4.png}. 
%
Costs  can be expressed
in the grammar as weights associated to each production rule.
For instance, the following production rules define two possible divisions
of a time interval into respectively a duplet and a triplet. Each comes with
a specific cost.
\[
\rho_1: q_0 \xrightarrow{0.06} \< q_{1}, q_{2}>,\
\rho_2: q_0 \xrightarrow{0.12} \< q_{1}, q_{2}, q_{2}>.
\]
Further binary divisions of time sub-intervals %represented respectively by~$q_2$ and~$q_3$
are possible with:
\[
\rho_3:\,q_2 \xrightarrow{0.1} \< q_{3}, q_{3}>, \
\rho_4:\,q_3 \xrightarrow{0.11} \< q_{4}, q_{4}>. \
\]

Score \includegraphics[scale=0.20]{pictures/score5.png} is obtained as follows:
the first measure
results from the successive applications of  rules $\rho_1$ (division in two)
and rules $\rho_3$ (division of the second half) in two. The second measure
is a division in three obtained by rule $\rho_3$. One obtains a parse tree
that can be linearized in the nested word of Example~\ref{ex:nested-word}.
...
\end{example}

@@

\begin{example}[Symbolic Weighted Parsing and the transcription problem]
Applied to the music transcription problem, the above formalism
is interpreted as follows:
%
\begin{itemize}
\item The input word is $I$ of Example~\ref{ex:running}.
\item The \SWT $T$ evaluates a ``fitness measure''  that expresses
    a correspondance between a performance and notation events. See Example~\ref{ex:SWT}.
\item The \SWVPA $A$ expresses a cost related to the music notation. 
\end{itemize}
One possibility is to relate this cost to the structural complexity: 
given a set of equivalent representations, 
we aim at choosing the simpler one. 
% 
For instance
\includegraphics[scale=0.20]{pictures/score5.png} will be favored
on \includegraphics[scale=0.20]{pictures/score4.png}. Costs  can be expressed
in the grammar as weights associated to each production rule.
For instance, the following production rules define two possible divisions
of a time interval into respectively a duplet and a triplet. Each comes with
a specific cost.
\[
\rho_1: q_0 \xrightarrow{0.06} \< q_{1}, q_{2}>,\
\rho_2: q_0 \xrightarrow{0.12} \< q_{1}, q_{2}, q_{2}>.
\]
Further binary divisions of time sub-intervals %represented respectively by~$q_2$ and~$q_3$
are possible with:
\[
\rho_3:\,q_2 \xrightarrow{0.1} \< q_{3}, q_{3}>, \
\rho_4:\,q_3 \xrightarrow{0.11} \< q_{4}, q_{4}>. \
\]

Score \includegraphics[scale=0.20]{pictures/score5.png} is obtained as follows:
the first measure
results from the successive applications of  rules $\rho_1$ (division in two)
and rules $\rho_3$ (division of the second half) in two. The second measure
is a division in three obtained by rule $\rho_3$. One obtains a parse tree
that can be linearized in the nested word of Example~\ref{ex:nested-word}.

Therefore, the framework, applied to the transcription problem, allows to find an optimal
solution that considers both the fitness of the result, and its structural complexity.
\end{example}

>>>>

\begin{example} \label{ex:SWVPA}
We consider a \SWVPA over the alphabet of Example~\ref{ex:nested-word}
that expresses a weight related to the music notation,
more precisely to its structural complexity: 
given a set of equivalent representations, we aim at choosing the simpler one. 
This si specified by the weights associated to the transition.

Let us assume a third component in the symbols of $\Deltac$ 
which is the duration of the associated time interval.
It is used to check the timestamps in the symbols.
%
For instance, the following call transition starts in state
$q_{{i}/{c}}$, meaning that the current interval is number~$i$ 
amongst $c$ sub-intervals of duration~$\ell$ (as indicated by the stack top).
It reads a new time-division symbol~$\ccall{d}$.
The date for this symbol has to be the time $\tau$ 
for the previous time-division symbol read~$\ccall{c}$ (found on the stack)
plus the total duration of intervals read so far, $i \ell$.
The new current interval duration is~$\frac{\ell}{d}$.

\(
\weic\bigl(q_{{i}/{c}}, 
           [\ccall{c}\ts{\tau}\ts{\ell}], q, 
           [\ccall{d}\ts{\tau + i\ell}\ts{\frac{\ell}{d}}], q_{{i+1}/{c}},
           q_{{0}/{d}} \bigr) = \alpha_d
\).

\noindent
The weight of this transition is $\alpha_d$. 
We can penalize triplets compared to duplets with $\alpha_2 < \alpha_3$.
%
\noindent 
Along with $\ccall{d}$, 
the above transition pushes
the state $q_{{i+1}/{c}}$ on the stack,  
to start the next sub-interval after reading $\creturn{d}$, 
with:
\(
\weir\bigl(q_{{d}/{d}}, 
           [\ccall{d}\ts{\tau}\ts{\ell}], q_{{i}/{c}}, 
           [\,\creturn{d}\ts{\tau + d\ell}], q_{{i}/{c}} \bigr) = \one
\).

\noindent
Reading an event is done with:
\(
\weii\bigl(q_{{i}/{d}}, 
           [\ccall{d}\ts{\tau}\ts{\ell}], q_{{j}/{c}}, 
           [\mu\ts{\tau + i\ell}], q_{{i+1}/{d}} \bigr) = \alpha_\mu
\).
% cas des barres de mesure?
\endex
\end{example}

@@

\begin{example}[Symbolic Weighted Parsing and the transcription problem]
Applied to the music transcription problem, the above formalism
is interpreted as follows:
%
\begin{itemize}
\item The input word is $I$ of Example~\ref{ex:running}.
\item The \SWT $T$ evaluates a ``fitness measure''  that expresses
    a correspondance between a performance and notation events. See Example~\ref{ex:SWT}.
\item The \SWVPA $A$ expresses a cost related to the music notation. 
\end{itemize}
%One possibility is to relate this cost to the structural complexity: 
%given a set of equivalent representations, 
%we aim at choosing the simpler one. 
% 
As seen in Example~\ref{ex:SWVPA}, 
\includegraphics[scale=0.35,trim=0 5mm 0 0]{pictures/ex1.pdf},
%\includegraphics[scale=0.20]{pictures/score5.png} 
will be favored on 
\includegraphics[scale=0.35,trim=0 5mm 0 0]{pictures/ex2.pdf}
%\includegraphics[scale=0.20]{pictures/score4.png}. 
when the weight a second time division with $\ccall{2}$ is less
than the difference of weight between 
$\mbox{`\footnotesize C$\sharp$5`}$
and $\mbox{\footnotesize C$\sharp$5}$.

\noindent
The SW-parsing framework, applied to the transcription problem, 
allows to find an optimal solution 
that considers both the fitness of the result, 
and its structural complexity.
\end{example}
